SHARP - II. Beware the effect of edge-on discs when interpreting the anomalous flux-ratios of gravitationally lensed quasars

### 2015-08-31 papers/B1555_EdgeOn/B1555_letter.tex
update of Modelling part

We use the lens modelling code {\sc gravlens} \citep{Kee01} to model the compact radio components. The inputs to the model are the observed image positions measured from the MERLIN radio observations of \citet{Marlow99} as all of the images are significantly detected in these data. The flux-ratio measurements are taken from \citet{K03}, who used half a year of MERLIN monitoring to obtain flux-ratio curves. In total, there are 11 constraints to the mass model provided by the observational data. As a first trial of modelling, we re-create the singular isothermal ellipsoid (SIE; 7 free parameters) model of \citet{Marlow99} to check the performance of a simple lensing potential. Consistent with previous studies \citep{Marlow99,Xu14}, a single elliptical potential model cannot reproduce the observed flux-ratios in B1555+375 (see Table~\ref{tab:results}; reduced $\chi^2 =~??$).

The next step is to examine if an edge-on disc can cause a physically plausible perturbation in the strong lensing mass model that results in the observed flux ratio anomaly of  merging images A and B. An exponential disc profile best describes the disc component in spiral galaxies because it matches the light distribution \citep{Kee98}. We thus choose a SIE plus an exponential disc model (SIE+ExpDisc) in our study of B1555+375. The free parameters of the model are the SIE Einstian radius ($b$), centroid position, ellipticity ($e$), and position angle ($\theta$), and the exponential disc intrinsic central density ($\kappa_0$), and scale length ($R_s$). A Detailed definition of these mass profiles is described by \citet{Kee01}. The source position and flux add three additional model parameters. We have not included a third component such as an external shear or an NFW halo because the number of constraints provided by the compact images detected by MERLIN is not sufficient to constrain such a model. Significantly deeper observations at a higher bandwidth with current very long baseline interferometric arrays will test more complex models in the future by imaging the extended gravitational arc and more significantly detecting image D.

To obtain the best-fit lens model, we explore the parameter space with the following steps. First, we apply the a priori information from the Keck AO image by setting the ellipticity to $e=0.8$ and the position angle to $\theta=10$ degr for the exponential disc. We then use the best-fit model of \citet{Marlow99} as the starting point for the SIE profile. The rest of the parameters that describe the exponential disc are set to be the same as the SIE component as the starting point. All of the model parameters are then allowed to vary, but we fix the centroid position of the SIE and exponential disc to the infrared positions with a Gaussian prior ($\sigma = 0.05 $~arcsec). The total number of free parameters is 11. After obtaining the best-fit lens model, we run a Markov-Chain Monte-Carlo (MCMC) analysis to explore the parameter space for the uncertainties.

The best-fit model parameters are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:model}. Table \ref{tab:results} also shows the comparison between the observed radio data and the model predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth potential model there is significant improvement in the predicted flux-ratios when compared to the results for the SIE-only model, with a reduced $\chi^2 = ~??$. The positions are all in agreement, and the flux-ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The mass model requires that the exponential disc component makes up about 10 per cent of the total mass within the Einstein radius (1.7~kpc for a lens redshift of $z_l = 1$), which is not unreasonable. We note that there is a significant 0.036~arcsec offset between the SIE and exponential disc components, which is likely due to there being an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models that use the full extent of the gravitational arc and the compact lensed images from future high resolution radio imaging will be able confirm this.

@@

\begin{table*}
\centering
\caption{A summary of the observational constraints and the predicted lens-model parameters. The lensed image positions in the radio \citep{Marlow99} (in arcsec) and average flux-ratios \citep{K03} are respect to image A. Uncertainties in the position measurements for images A, B and C are $0.001$~arcsec and for image D are $0.006$~arcsec.}
\begin{tabular}{cccccccc}
\hline
Image	&\multicolumn{3}{c}{Observed MERLIN} 	 	& \multicolumn{4}{c}{Model-Predicted}\\
		&\multicolumn{3}{c}{5 GHz}		& & & {SIE+ExpDisc} & SIE-only\\
		&East &North & Flux ratio &East 	&North & Flux ratio &Flux ratio\\ 
\hline
A  &$0$    		&$0$			&  $1$ 				&$+0.000$  &$+0.000$	& $1$ 		& $1$\\  
B  &$-0.073$	&$+0.048$	& $0.620 \pm 0.039$ 	&$-0.073$ &$+0.048$		& $0.620$ 	& $0.962$ \\  
C  &$-0.412$ 	&$-0.028$	& $0.507\pm 0.030$	&$-0.412$ &$-0.028$		& $0.461$ 	& $0.313$\\  
D  &$-0.162$	&$-0.368$	& $0.086 \pm 0.024$ 	&$-0.163$ &$-0.365$		& $0.115$ 	& $0.103$\\  
\hline
\end{tabular}
\label{tab:results}
\end{table*}

\begin{table}
\centering
\caption{Best-fit model parameters of B1555+375 for an SIE + exponential disc lens model and the uncertainties. The positions are measured in arcsec, relative to the radio position of image A. The ellipticity is defined as the axis ratio $q=1-e$ and the position angles are measured in degrees east of north. The Einstein radius $b$ and disc-scale radius are measured in arcsec.}
\begin{tabular}{ccc}
\hline 
Parameter  & SIE & ExpDisc \\
\hline
$\Delta$RA	& $-0.182 \pm 0.005$		& $-0.147 \pm 0.005$\\
$\Delta$Dec	& $-0.199 \pm 0.005$		& $-0.206 \pm 0.006$ \\
$b$ 			& $0.21 \pm 0.01$  		& $0.11 \pm 0.01$ \\
$e$	  		& $0.31 \pm 0.01$			& $0.87 \pm 0.01$ \\
$\theta$ 		& $3.0 \pm 0.6$			& $7.0 \pm 0.7$	 \\
$R_s$			& ...  						& $0.28 \pm 0.01$	 \\
\hline
\label{tab:model}
\end{tabular}
\end{table}

>>>>

We use the lens modelling code {\sc gravlens} \citep{Kee01} to model the compact radio components. The inputs to the model are the observed image positions measured from the MERLIN radio observations of \citet{Marlow99} as all of the images are significantly detected in these data. The flux-ratio measurements are taken from \citet{K03}, who used half a year of MERLIN monitoring to obtain flux-ratio curves. In total, there are 11 constraints to the mass model provided by the observational data. As a first trial of modelling, we re-create the singular isothermal ellipsoid (SIE; 7 free parameters) model of \citet{Marlow99} to check the performance of a simple lensing potential. Consistent with previous studies \citep{Marlow99,Xu14}, a single elliptical potential model cannot reproduce the observed flux-ratios in B1555+375 (see Table~\ref{tab:results}; reduced $\chi^2 =97.3$).

The next step is to examine if an edge-on disc can cause a physically plausible perturbation in the strong lensing mass model that results in the observed flux ratio anomaly of  merging images A and B. An exponential disc profile best describes the disc component in spiral galaxies because it matches the light distribution \citep{Kee98}. We thus choose a SIE plus an exponential disc model (SIE+ExpDisc) in our study of B1555+375. The free parameters of the model are the SIE Einstien radius ($b$), centroid position, ellipticity ($e$), position angle ($\theta$), the exponential disc intrinsic central density ($\kappa_0$), and scale length ($R_s$). A Detailed definition of these mass profiles is described by \citet{Kee01}. The source position add two additional model parameters. We have not included a third component such as an external shear or an NFW halo because the number of constraints provided by the compact images detected by MERLIN is not sufficient to constrain such a model. Significantly deeper observations at a higher bandwidth with current very long baseline interferometric arrays will test more complex models in the future by imaging the extended gravitational arc and more significantly detecting image D.

To obtain the best-fit lens model, we explore the parameter space with the following steps. First, we apply the a priori information from the Keck AO image by setting the ellipticity to $e=0.8$ and the position angle to $\theta=10$ degr for the exponential disc. We then use the best-fit model of \citet{Marlow99} as the starting point for the SIE profile. The rest of the parameters that describe the exponential disc are set to be the same as the SIE component as the starting point. All of the model parameters are then allowed to vary, but we fix the centroid position of the SIE and exponential disc to the infrared positions with a Gaussian prior ($\sigma = 0.05 $~arcsec). The total number of free parameters is 13. After obtaining the best-fit lens model, we run a Markov-Chain Monte-Carlo (MCMC) analysis to explore the parameter space for the uncertainties.

The best-fit model parameters are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:model}. Table \ref{tab:results} also shows the comparison between the observed radio data and the model predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth potential model there is significant improvement in the predicted flux-ratios when compared to the results for the SIE-only model, with a reduced $\chi^2 = 2.7$. The positions are all in agreement, and the flux-ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The mass model requires that the exponential disc component makes up about 10 per cent of the total mass within the Einstein radius (1.7~kpc for a lens redshift of $z_l = 1$), which is not unreasonable. We note that there is a significant 0.036~arcsec offset between the SIE and exponential disc components, which is likely due to there being an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models that use the full extent of the gravitational arc and the compact lensed images from future high resolution radio imaging will be able confirm this.

@@

\begin{table*}
\centering
\caption{A summary of the observational constraints and the predicted lens-model parameters. The lensed image positions in the radio \citep{Marlow99} (in arcsec) and average flux-ratios \citep{K03} are respect to image A. Uncertainties in the position measurements for images A, B and C are $0.001$~arcsec and for image D are $0.006$~arcsec.}
\begin{tabular}{cccccccc}
\hline
Image	&\multicolumn{3}{c}{Observed MERLIN} 	 	& \multicolumn{4}{c}{Model-Predicted}\\
		&\multicolumn{3}{c}{5 GHz}		& & & {SIE+ExpDisc} & SIE-only\\
		&East &North & Flux ratio &East 	&North & Flux ratio &Flux ratio\\ 
\hline
A  &$0$    		&$0$			&  $1$ 				&$+0.000$  &$+0.000$	& $1$ 		& $1$\\  
B  &$-0.073$	&$+0.048$	& $0.620 \pm 0.039$ 	&$-0.073$ &$+0.048$		& $0.620$ 	& $0.971$ \\  
C  &$-0.412$ 	&$-0.028$	& $0.507\pm 0.030$	&$-0.412$ &$-0.028$		& $0.461$ 	& $0.312$\\  
D  &$-0.162$	&$-0.368$	& $0.086 \pm 0.024$ 	&$-0.163$ &$-0.365$		& $0.115$ 	& $0.106$\\  
\hline
\end{tabular}
\label{tab:results}
\end{table*}

\begin{table}
\centering
\caption{Best-fit model parameters of B1555+375 for an SIE + exponential disc lens model and the uncertainties. The positions are measured in arcsec, relative to the radio position of image A. The ellipticity is defined as the axis ratio $q=1-e$ and the position angles are measured in degrees east of north. The Einstein radius $b$ and disc-scale radius are measured in arcsec.}
\begin{tabular}{ccc}
\hline 
Parameter    & Comp. 1  & Comp. 2  \\
   & (SIE) & (ExpDisc)\\
\hline
$\Delta$RA	& $-0.182 \pm 0.005$		& $-0.147 \pm 0.005$\\
$\Delta$Dec	& $-0.199 \pm 0.005$		& $-0.206 \pm 0.006$ \\
$b$ 			& $0.21 \pm 0.01$  		& $0.11 \pm 0.01$  \\
$e$	  		& $0.31 \pm 0.01$			& $0.87 \pm 0.01$ \\
$\theta$ 		& $3.0 \pm 0.6$			& $7.0 \pm 0.7$	 \\
$R_s$			& ...  						& $0.28 \pm 0.01$	 \\
\hline
\label{tab:model}
\end{tabular}
\end{table}
